### 2019-06-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
OMG c'est 00h23 un dimanche soir et je bosse !

Assume that we collect independent samples $X_1,X_2,\dots$ all from some distributions supported in $[0,1]$.
We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\dots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\dots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall i\in\N^*, \bE[X_i] = \mu_0\right)$.

@@

Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows before hand the values of $a$ and $b$).

@@

If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem introduced above would reduce to a parametric sequential test of
$\cH_0 : (\exists \mu_0: \forall i\in\N, X_i \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
against the alternative
$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.
%
The Generalized Likelihood Ratio statistic for this test is defined by
\begin{equation}\label{eq:6:firstDefGLRT}
    \GLR(n) \eqdef \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)},
\end{equation}
%
where $\ell(X_1, \dots, X_n ; \mu_0)$ and $\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the \emph{likelihoods} of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$),
$\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
%
Indeed, we show below
% in Appendix~\ref{app:6:GLR_with_kl}
that in any one-dimensional exponential family, we have
\begin{equation}\label{eq:6:secondDefGLRT}
    \log \GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right].
\end{equation}
%
where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{l=k}^{k'} X_l$ denotes the average of the observations $X_l$ collected between the instants $k$ and $k'$.

\begin{smallproof}
    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod_{i=1}^n \ell(X_k ; \mu_0)$ by independence of the observations $X_k$,
    and $\ell(X_k ; \mu_0) = \mu_0^{X_k} (1-\mu_0)^{1-X_k}$ for Bernoulli distributions.
    Therefore, taking the logarithm gives
    \begin{align*}
        \log\ell(X_1, \dots, X_n ; \mu_0)
        &= \sum_{i=1}^n X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) \\
        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_k \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_k \right) \\
        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
    \end{align*}
    %
    For a constant $a\in[0,1]$, let $h(x) \eqdef a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative:
    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
    %
    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
        \sup_{\mu_0} \log\ell(X_1, \dots, X_n ; \mu_0) = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
    \end{equation}

@@

        & \log\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \\
        &= \sum_{i=1}^{\tau} X_k \log(\mu_0) + (1-X_k) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_k \log(\mu_1) + (1-X_k) \log(1-\mu_1) \\

@@

    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_k = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,

@@

\begin{lemma}\label{lem:6:FalseAlarm}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_k \in [0,1]$ for all $i$.
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    $\bP_{\mu_0}(\widehat{\tau}_\delta < \infty) \leq \delta$, with the threshold function
    %
    \begin{equation}\label{def:6:beta}
        \beta(n,\delta) = 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

\begin{smallproof}
Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).

Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
We first observe that
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists (s,n) \in \N^2, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
% \end{align*}%FIXME uncomment this if needed
% \begin{align*}
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^*:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6F(s+r)+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3F(s)+3F(r) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
%
\end{smallproof}

\begin{lemma}\label{lem:6:ConcFirst}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    $(X_k)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ two independent \iid{} processes with respective means $\mu$ and $\mu'$ such that
    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$,
    where $\phi_\mu(\lambda) = \bE_{X \sim \nu^{\mu}}[e^{\lambda X}]$ is the moment generating function of the distribution $\nu^\mu$, which is the unique distribution in an exponential family that has mean $\mu$.

    Let $\kl(\mu,\mu') = \KL(\nu^\mu,\nu^{\mu'})$ be the divergence function associated to that exponential family. Introducing the notation $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_k$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$, it holds that for every $s,r \in \N^*$,
    \[\bP\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu\right) + r \, \kl\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},\]
    where $\cT$ is the function defined in \eqref{def:6:function_T}.
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
\begin{smallproof}
    This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
\end{smallproof}

@@

\begin{lemma}\label{lem:6:Delay}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
    and let $\Delta = |\mu_0 - \mu_1|$.
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    \begin{equation}
        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp\left( -\frac{2\tau u}{\tau + u}\left(\max\left[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} \right]\right)^2 \right).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
%
\begin{smallproof}
    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
    % \TODOL{We have to give the proof, or remove this result? Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
\end{smallproof}

@@

Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible positions for a change-point. Thus, in practice the following adaptation may be interesting, based on down-sampling the possible time steps:
%
\begin{equation}\label{def:6:GLRTricks}
    \widetilde{\tau}_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
\end{equation}
%
for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the same false-alarm control.
However, the detection delay may be slightly increased.
In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups
has little impact on the regret of the bandit strategy introduced below.

@@

\begin{proposition}\label{prop:6:EnoughSamples}
\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$ (\ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$) it holds that
    \begin{equation}
        n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor.
    \end{equation}
\end{leftbar}  % XXX leftbar propositionbar, comment if needed
\end{proposition}

@@

This section give the finite-time regret bounds for the two variants, and interpretations of the different results.

@@

One can note that the best non-oracle strategies are actively adaptive,
thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.

>>>>

Assume that we collect independent samples $X_1,X_2,\dots$ all from some distributions supported in $[0,1]$.
We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\dots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\dots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall t\in\N^*, \bE[X_t] = \mu_0\right)$.

@@

Lemma~1 of \cite{KLUCBJournal} establishes that any bounded distribution supported in $[0,1]$ is a sub-Bernoulli distribution.
As explained in Section~\ref{par:2:normalizationTrick}, our work can be applied to any bounded distribution without loss of generality, as a reward in $r\in[a,b]$ can clearly be mapped to $[0,1]$ simply by using $r' = (r-a)/(b-a)$ (we assume that the decision maker or the algorithm knows beforehand the values of $a$ and $b$).

@@

If the samples $(X_t)$ were all drawn from a Bernoulli distribution, this change-point detection problem would reduce to a parametric sequential test of
$\cH_0 : (\exists \mu_0: \forall t\in\N^*, X_t \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0))$,
against the alternative
$\cH_1 : (\exists \mu_0 \neq \mu_1, \tau \in \N^*: \ \  X_1, \dots, X_\tau \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_0) \ \ \text{and} \ \ X_{\tau+1}, X_{\tau+2}, \dots \ \overset{\text{i.i.d.}}{\sim} \ \cB(\mu_1))$.\\
%
The Generalized Likelihood Ratio statistic for this test is defined by
\begin{equation}\label{eq:6:firstDefGLRT}
    \GLR(n) \eqdef \frac{\sup\limits_{\mu_0,\mu_1,\tau < n}\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)}{\sup\limits_{\mu_0}\ell(X_1, \dots, X_n ; \mu_0)},
\end{equation}
%
where $\ell(X_1, \dots, X_n ; \mu_0)$ and $\ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau)$ respectively denote the \emph{likelihoods} of the first $n$ observations under a model in $\cH_0$ and $\cH_1$.
High values of this statistic tend to indicate rejection of $\cH_0$.
%
By using the form of the likelihood for Bernoulli distributions, this statistic can be written with the binary relative entropy $\kl$, defined for $x,y\in[0,1]$ (with the usual convention that $t \log(t) = 0$ if $t=0$),
$\kl(x,y) \eqdef x \ln\left(\frac{x}{y}\right) + (1-x)\ln\left(\frac{1-x}{1-y}\right)$
%
Indeed, we show below
% in Appendix~\ref{app:6:GLR_with_kl}
that for Bernoulli distributions, we have
\begin{equation}\label{eq:6:secondDefGLRT}
    \log \GLR(n) = \sup_{s \in [2,n-1]} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right].
\end{equation}
%
where for $k \leq k'$, $\widehat{\mu}_{k:k'} \eqdef \frac{1}{k - k' + 1} \sum\limits_{t=k}^{k'} X_t$ denotes the average of the observations $X_t$ collected between the instants $k$ and $k'$.

\begin{smallproof}
    First, we consider the denominator in the expression of $\GLR(n)$ \eqref{eq:6:firstDefGLRT}, that is the $\sup$ on $\mu_0$.
    We have $\ell(X_1, \dots, X_n ; \mu_0) = \prod\limits_{i=1}^n \ell(X_i ; \mu_0)$ by independence of the observations $X_i$,
    and $\ell(X_i ; \mu_0) = \mu_0^{X_i} (1-\mu_0)^{1-X_i}$ for Bernoulli distributions.
    Therefore, taking the logarithm gives
    \begin{align*}
        \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right]
        &= \sum_{i=1}^n X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) \\
        &= \log(\mu_0) \times \left( \sum_{i=1}^n X_i \right) + \log(1-\mu_0) \times \left( n - \sum_{i=1}^n X_i \right) \\
        &= n \left( \widehat{\mu}_{1:n} \log(\mu_0) + (1 - \widehat{\mu}_{1:n}) \log(1-\mu_0) \right).
    \end{align*}
    %
    For a constant $a\in[0,1]$, let $h(x) \eqdef a \log(x) + (1-a) \log(1-x)$ on $(0,1)$,
    and we are trying to solve $\sup_{x\in[0,1]} h(x)$.
    If $a=0$ or $a=1$, $h$ is maximum at $x=a$.
    Now if $a\neq0$, As $h$ is of class $\cC^1$, we can just differentiate and find the root of its derivative,
    $h'(x) = a/x - (1-a)/(1-x)$, so $h'(x) = 0$ if and only if $x=a$.
    Thus in all cases, $\sup_{x\in[0,1]} h(x) = h(a)$.
    %
    Here, we have $a = \widehat{\mu}_{1:n}$, and thus we solve the $\sup_{\mu_0}$ optimization problem found in the denominator of the $\GLR(n)$ expression.
    By replacing $\mu_0 = \widehat{\mu}_{1:n}$, we obtained the following (unique) solution,
    \begin{equation}\label{eq:6:solutionPbOpt_for_1n}
        \sup_{\mu_0} \log \left[ \ell(X_1, \dots, X_n ; \mu_0) \right] = n \left( \widehat{\mu}_{1:n} \log(\widehat{\mu}_{1:n}) + (1 - \widehat{\mu}_{1:n}) \log(1-\widehat{\mu}_{1:n}) \right).
    \end{equation}

@@

        & \log \left[ \ell(X_1, \dots, X_n ; \mu_0,\mu_1,\tau) \right] \\
        &= \sum_{i=1}^{\tau} X_i \log(\mu_0) + (1-X_i) \log(1-\mu_0) + \sum_{i=\tau+1}^n X_i \log(\mu_1) + (1-X_i) \log(1-\mu_1) \\

@@

    By definition, $n \; \widehat{\mu}_{1:n} = \sum_{i=1}^n X_i = s \;\widehat{\mu}_{1:s} + (n-s) \;\widehat{\mu}_{s+1:n}$, so the right hand side (negative) part involving $\widehat{\mu}_{1:n}$ can be distributed in the two left hand side (positive) terms,

@@

\begin{lemma}\label{lem:6:FalseAlarm}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t$.
    % Assume that there exists $\mu_0 \in [0,1]$ such that $\bE[X_t] = \mu_0$ and that $X_i \in [0,1]$ for all $i$.
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    $\bP_{\mu_0}(\widehat{\tau}_\delta < \infty) \leq \delta$, with the threshold function
    %
    \begin{equation}\label{def:6:beta}
        \beta(n,\delta) = 2\cT\left(\frac{\ln(3n\sqrt{n}/\delta)}{2}\right) + 6\ln(1+\ln(n)).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

\begin{smallproof}
Lemma~\ref{lem:6:FalseAlarm} is presented for bounded distributions and is actually valid for any sub-Bernoulli distribution. It could also be presented for more general distributions satisfying
\begin{equation}
    \bE[e^{\lambda X} ] \leq e^{\phi_\mu(\lambda)} \ \ \text{ with } \ \ \mu=\bE[X]\label{def:6:subExpo},
\end{equation}
where $\phi_\mu(\lambda)$ is the $\mathrm{log}$ moment generating of some one-dimensional exponential family. The Bernoulli divergence $\kl(x,y)$ would be replaced by the corresponding divergence in that exponential family (which is the Kullback-Leibler divergence between two distributions of means $x$ and $y$).
%
Let us go back to the Bernoulli case with divergence $\kl(x,y)$.
We first have
\[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) = \inf_{\lambda \in [0,1]}\left[s \times \kl\left(\widehat{\mu}_{1:s},\lambda\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\lambda\right)\right].\]
Hence the probability of a false alarm occurring is upper bounded as
%
\begin{align*}
    \bP_{\mu_0}\left(T_{\delta} < \infty\right) & \leq \bP_{\mu_0}\left(\exists s \in \N^*, n \in \N^*, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right) > \beta(n,\delta)\right)\\
    & \leq \bP_{\mu_0}\left(\exists s \in \N^*, n \in \N^*, s < n:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right) \\
% \end{align*}%FIXME uncomment this if needed
% \begin{align*}
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists n > s:  s \, \kl\left(\widehat{\mu}_{1:s},\mu_0\right) + (n-s) \, \kl\left(\widehat{\mu}_{s+1:n},\mu_0\right) > \beta(n,\delta)\right)\\
    & =   \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^*:  s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > \beta(s+r,\delta)\right),
\end{align*}
%
where $\widehat{\mu}_{s}$ and $\widehat{\mu}'_{r}$ are the empirical means of respectively $s$ and $r$ \iid{} observations with mean $\mu_0$ and distribution $\nu$, that are independent from the previous ones.
As $\nu$ is sub-Bernoulli, the conclusion follows from Lemma~\ref{lem:6:ConcFirst} below and from the definition of $\beta(n,\delta)$, if we denote $F(x)=\ln(1+\ln(x))$,
%
\begin{align*}
    &\bP_{\mu_0}\left(T_{\delta} < \infty\right) \\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 6F(s+r)+ 2 \cT\left(\frac{\ln(3(s+r)^{3/2}/\delta)}{2}\right)\right)\\
    & \leq \sum_{s =1}^{\infty} \bP_{\mu_0}\left(\exists r \in \N^* : s \, \kl\left(\widehat{\mu}_{s},\mu_0\right) + r \, \kl\left(\widehat{\mu}'_{r},\mu_0\right) > 3F(s)+3F(r) + 2 \cT\left(\frac{\ln(3s^{3/2}/\delta)}{2}\right)\right)
\end{align*}
%
And so we have $\bP_{\mu_0}\left(T_{\delta} < \infty\right) \leq \sum_{s =1}^{\infty} \frac{\delta}{3s^{3/2}} \leq \delta$.
%
\end{smallproof}

\begin{lemma}\label{lem:6:ConcFirst}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Consider a one-dimensional exponential family $\cE$,
    and let $\nu_{\mu}$ be the unique distribution in this family that has mean $\mu$,
    with moment generating function $\phi_\mu(\lambda) = \bE_{X \sim \nu_{\mu}}[e^{\lambda X}]$.
    Let $\kl_{\cE}(\mu,\mu') = \KL(\nu_{\mu},\nu_{\mu'})$ be the divergence function associated to $\cE$.
    Let $(X_i)_{i \in \N^*}$ and $(Y_k)_{i \in \N^*}$ be two independent \iid{} processes, with respective means $\mu$ and $\mu'$, such that
    $\bE[e^{\lambda X_1}] \leq e^{\phi_{\mu}(\lambda)}$ and $\bE[e^{\lambda Y_1}] \leq e^{\phi_{\mu'}(\lambda)}$.
    %
    Denote $\widehat{\mu}_s = \frac{1}{s}\sum\limits_{i=1}^s X_i$ and $\widehat{\mu}'_r = \frac{1}{r}\sum\limits_{i=1}^r Y_k$.
    Then for every $s,r \in \N^*$ we have,
    \begin{small}
        \begin{equation}
            \bP\left(\exists r \in \N^* : s \, \kl_{\cE}\left(\widehat{\mu}_{s},\mu\right) + r \, \kl_{\cE}\left(\widehat{\mu}_r',\mu'\right) > 3\ln(1+\ln(s)) + 3\ln(1+\ln(r)) + 2\cT\left(\frac{x}{2}\right)\right)\leq e^{-x},
        \end{equation}
    \end{small}
    where $\cT$ is the function defined in \eqref{def:6:function_T}.
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
\begin{smallproof}
    This Lemma~\ref{lem:6:ConcFirst} is proved in Appendix~\ref{app:6:proofConcFirst}.
\end{smallproof}

@@

\begin{lemma}\label{lem:6:Delay}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    Let be $\bP_{\mu_0,\mu_1,\tau}$ a probabilistic model under which $X_t \in [0,1]$, and $X_t$ has an average of $\mu_0$ for all $t \leq \tau$, and $\mu_1$ for all $t > \tau$, with $\mu_0 \neq \mu_1$,
    and let $\Delta = |\mu_0 - \mu_1|$.\\
    Then the B-GLRT test given in Definition~\ref{def:6:GLRDef} satisfies
    \begin{equation}
        \bP_{\mu_0,\mu_1,\tau} (\widehat{\tau}_{\delta} \geq \tau + u) \leq \exp\left( -\frac{2\tau u}{\tau + u}\left(\max\left[ 0, \Delta - \sqrt{\frac{\tau + u}{2\tau u} \beta(\tau + u,\delta)} \right]\right)^2 \right).
    \end{equation}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}
%
\begin{smallproof}
    It is actually not used as such in the proofs of the regret upper bounds given below, because a similar but more specific result is used and proved in the proofs.
    For instance, see below in Section~\ref{par:6:controllingDelayInOneProof} in page~\pageref{par:6:controllingDelayInOneProof}.
    % \TODOL{We have to give the proof, or remove this result? Je ne vois pas le problème de le laisser là, ça formalise le résultat mais on a pas besoin de celui là, ça permet de bien formaliser ce qu'on entend par contrôler le délai, et dans la preuve plus bas on prouve ce qu'on souhaite.}
\end{smallproof}

@@

Also note that, as any test based on scan-statistics, the B-GLRT can be costly to implement as at every time step, it considers all previous time steps as a possible position for a change-point.
Thus, down-sampling the possible time steps is an interesting adaptation in practice:
%
\begin{equation}\label{def:6:GLRTricks}
    \widetilde{\tau}_{\delta} = \inf \left\{ n \in \cN : \sup_{s \in \cS_n} \left[s \times \kl\left(\widehat{\mu}_{1:s},\widehat{\mu}_{1:n}\right) + (n-s) \times \kl\left(\widehat{\mu}_{s+1:n},\widehat{\mu}_{1:n}\right)\right] \geq \beta(n,\delta)\right\},
\end{equation}
%
for subsets $\cN$ and $\cS_n$. Following the proof of Lemma~\ref{lem:6:FalseAlarm}, we can easily see that this variant enjoys the same false-alarm control.
However, the detection delay may be slightly increased.
In Appendix~\ref{sub:6:IdeasOptimizations} we show that using these practical speedups
has little impact on the regret of the bandit strategy introduced below.

@@

\begin{proposition}\label{prop:6:EnoughSamples}
\begin{leftbar}[propositionbar]  % XXX leftbar propositionbar, comment if needed
    For every pair of instants $s \leq t\in\N^*$ between two restarts on arm $k$ (\ie, for a $i\in[\NCk]$, one has $\tau_k^{(i)} = \tau_k(t) < s \leq t < \tau_k^{(i+1)}$) it holds that
    % \begin{equation}
    $n_k(t) - n_k(s) \geq \left\lfloor \frac{\alpha}{K} (t-s) \right\rfloor$.
    % \end{equation}
\end{leftbar}  % XXX leftbar propositionbar, comment if needed
\end{proposition}

@@

This section give the finite-time regret bounds for the two variants, and interpretations of the different results.
Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.

@@

We highlight that the best non-oracle strategies are actively adaptive,
thus the experiments confirm that an efficient bandit algorithm (\eg, \klUCB) combined with an efficient change point detector (\eg, \GLR) provides efficient strategies for the piece-wise stationary model.
